Switch equivalence and weak prime domains for fusions

### 2023-09-15 fusione.tex
cose

\title{Fusion}

@@

\begin{lemma}\cite{heindel2009category}
	Let $\X$ be an $\mathcal{M}, \mathcal{N}$-adhesive category and $(\X, R)$ a left-linear DPO-rewriting system over it. Then for every non-empty derivation $\der{D}:G\to H$, the diagram $\Delta(\der{D})$ has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ such that $\iota_H$ belong to $\mathcal{M}$.
\end{lemma}
\begin{proof}
	We proceed by induction.
	\begin{enumerate}
		\item Suppose that $n=0$ so that $\der{D}$ consists only of the following direct derivation
		\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G & \ar[l]^{f} D \ar[r]_{g}& H }\]
		then $f\in \mathcal{M}$ so that we can consider its pushout along $g$ to get a diagram
		\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G \ar@/_.2cm/[dr]_{q}& \ar[l]^{f} D \ar[r]_{g}& H  \ar@/^.2cm/[dl]^{p}\\ & P}\]
		Let $(C, \{c_X\}_{X\in \Delta(\der{D})})$ be another cocone, in particular we have
		\begin{align*}
			c_H\circ g & = c_D \\&= c_G\circ f
		\end{align*}
		Hence, there exists a unique $c:P\to C$ such that 
		\[c\circ q= c_G \qquad c\circ p =c_H\]
		Now, $f$ is in $\mathcal{M}$ as it is the pushout of $l$, hence $p$ is in $\mathcal{M}$ too and this gives us the thesis. 
		\item Let $\der{D}$ be $\{\dder{D}_i\}_{i=0}^n$ and suppose that $n\geq 1$. Let also $\der{D}'$ be $\{\dder{D}_i\}^{n-1}_{i=0}$ and $\rho_n=(l_n, r_n)$ be the rule applied in $\dder{D}_n$. The pushout of $l_n$ is the arrow $f_n:D_n\to G_n$ is in $\mathcal{M}$ and, by inductive hypothesis, $\iota_{G_{n}}:G_{n}\to \lpro \der{D}'\rpro$ is in $\mathcal{M}$ too. Thus, we can consider the diagram below, having a pushout as its lower half.
		\[\xymatrix{L_n \ar[d]_{m_{n}}& K_{n} \ar[d]^{k_{n}}\ar[l]_{l_{n}} \ar[r]^{r_{n}} & R_{n} \ar[d]^{h_n} \\G_{n} \ar[d]_{\iota_{G_{n}}}& \ar[l]^{f_n} D_n \ar[r]_{g_n}& H  \ar[d]^{q}\\ \lpro \der{D}' \ar[rr]_{p}\rpro && P}\] 
		Notice that, as in the point above, the arrow $q:H\to P$ is the pushout of an element in $\mathcal{M}$, therefore it is enough to show that the diagram so constructed provides a colimiting cocone for $\Delta(\der{D})$.

@@

\begin{definition}
	Let $\X$ be an $\mathcal{M}$-adhesive category and consider a left-linear DPO-rewriting system $(\X, R)$ on it.  Take two non-empty derivations $\der{D}=\{\dder{D}_i\}_{i=0}^n$ and $\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ with the same length with their associated sequences of rules $r(\der{D})=\{\rho_i\}_{i=0}^n$ and $r(\der{D}')=\{\rho'_i\}_{i=0}^n$. A permutation $\sigma:n+1\to n+1$  is called a \emph{consistent permutation} between $\der{D}$ and $\der{D}'$, if for every $i\in n+1$, $\rho_i=\rho'_{\sigma(i)}$ and, moreover, there exists an isomorphism $\phi:\tpro{D} \to \lpro \der{D}' \rpro$ fitting in the following diagrams, where $m_i, m'_i, h_i$ and $h'_i$ are, respectively, the matches and back-matches of $\dder{D}_i$ and $\dder{D}'_i$.
\[\xymatrix{L_i \ar[r]^{m_i} \ar[d]_{m'_{\sigma(i)}}& G_i \ar[r]^{\iota_{G_i}} &\tpro{D} \ar[d]^{\xi} & R_i \ar[r]^{h_i} \ar[d]_{h'_{\sigma(i)}}& G_{i+1} \ar[r]^{\iota_{G_{i+1}}} &\tpro{D} \ar[d]^{\xi} \\G'_{\sigma(i)} \ar[rr]_{\iota_{G'_{\sigma(i)}}}&& \lpro \der{D}' \rpro& G'_{\sigma(i)+1} \ar[rr]_{\iota_{G'_{\sigma(i)+1}}}&& \lpro \der{D}' \rpro}\]
\end{definition}

@@

Take two derivations $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and $\der{D}'=\{\dder{D}'_{i}\}_{i=0}^n$ with the same length and between the same $G_0$ and $G_n$. We say that $\der{D}'$ is \emph{obtained by a proper switch from $\der{D}$} if there exists an index $j < n$ such that
	\begin{enumerate}
\item for every $i\notin \{j, j+1\} $, $\dder{D}_i=\dder{D}'_i$;
\item $\dder{D}_j \Updownarrow \dder{D}_{j+1}$;		
\item $\dder{D}'_j\cdot \dder{D}'_{j+1} = S_{u,u'}(\dder{D}, \dder{D}')$.
	\end{enumerate}

	We will also say that $\der{D}$ is \emph{switch equivalent} to $\der{D}'$, if there exists a sequence, $\{\der{D}_i\}_{i=0}^n$ of derivations such that
	\begin{enumerate}
		\item $\der{D}_0=\der{D}$ and $\der{D}_n=\der{D}'$;
		\item for every $i< n$, $\der{D}_{i+1}$ is obtained by a proper switch from $\der{D}_i$.
	\end{enumerate}

@@

\begin{proposition}
	\todo{reversibilità}
\end{proposition}
\begin{proof}
	contenuto...
\end{proof}
\begin{corollary}
	\todo{ho davver una relazione di equivalenza}
\end{corollary}

@@

\begin{lemma}\todo{Punto 1 del Lemma 17}
	contenuto...
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}

>>>>

\title{Switch equivalence and weak prime domains for fusions}

@@

\begin{lemma}\label{lem:colim}
	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, R)$ a left-linear DPO-rewriting system over it. The following hold true:
	\begin{enumerate}
		\item if $\der{D}$ is a derivation of length one given by the following diagram

		then the colimiting cocone $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ exists and  fits in the diagram below, wher the bottom face is a pushout;
		\[\xymatrix{}\]
		\item 
		\item  for every non-empty derivation $\der{D}:G\to H$, the diagram $\Delta(\der{D})$ has a colimit $(\tpro{D}, \{\iota_X\}_{X\in \Delta(\der{D})})$ such that $\iota_H$ belong to $\mathcal{M}$.
	\end{enumerate}
\end{lemma}
\begin{proof}\begin{enumerate}
		contenuto...
	\end{enumerate}
	contenuto...
\end{proof}

\begin{lemma}	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, R)$ a left-linear DPO-rewriting system over it.
 Then
\end{lemma}
\begin{proof}
	We proceed by induction.
	\begin{enumerate}
		\item Suppose that $n=0$ so that $\der{D}$ consists only of the following direct derivation
		\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G & \ar[l]^{f} D \ar[r]_{g}& H }\]
		then $f\in \mathcal{M}$ so that we can consider its pushout along $g$ to get a diagram
		\[\xymatrix{L \ar[d]_{n}& K \ar[d]^{k}\ar[l]_{l} \ar[r]^{r} & R \ar[d]^{h} \\G \ar@/_.2cm/[dr]_{q}& \ar[l]^{f} D \ar[r]_{g}& H  \ar@/^.2cm/[dl]^{p}\\ & P}\]
		Let $(C, \{c_X\}_{X\in \Delta(\der{D})})$ be another cocone, in particular we have
		\begin{align*}
			c_H\circ g & = c_D \\&= c_G\circ f
		\end{align*}
		Hence, there exists a unique $c:P\to C$ such that 
		\[c\circ q= c_G \qquad c\circ p =c_H\]
		Now, $f$ is in $\mathcal{M}$ as it is the pushout of $l$, hence $p$ is in $\mathcal{M}$ too and this gives us the thesis. 
		\item Let $\der{D}$ be $\{\dder{D}_i\}_{i=0}^n$ and suppose that $n\geq 1$. Let also $\der{D}'$ be $\{\dder{D}_i\}^{n-1}_{i=0}$ and $\rho_n=(l_n, r_n)$ be the rule applied in $\dder{D}_n$. The pushout of $l_n$ is the arrow $f_n:D_n\to G_n$ is in $\mathcal{M}$ and, by inductive hypothesis, $\iota_{G_{n}}:G_{n}\to \lpro \der{D}'\rpro$ is in $\mathcal{M}$ too. Thus, we can consider the diagram below, having a pushout as its lower half.
		\[\xymatrix{L_n \ar[d]_{m_{n}}& K_{n} \ar[d]^{k_{n}}\ar[l]_{l_{n}} \ar[r]^{r_{n}} & R_{n} \ar[d]^{h_n} \\G_{n} \ar[d]_{\iota_{G_{n}}}& \ar[l]^{f_n} D_n \ar[r]_{g_n}& H  \ar[d]^{q}\\ \lpro \der{D}' \ar[rr]_{p}\rpro && P}\] 
		Notice that, as in the point above, the arrow $q:H\to P$ is the pushout of an element in $\mathcal{M}$, therefore it is enough to show that the diagram so constructed provides a colimiting cocone for $\Delta(\der{D})$.

@@

\begin{definition}
	Let $\X$ be an $\mathcal{M}$-adhesive category and consider a left-linear DPO-rewriting system $(\X, R)$ on it.  Take two non-empty derivations $\der{D}=\{\dder{D}_i\}_{i=0}^n$ and $\der{D}'=\{\dder{D}'_i\}_{i=0}^n$ with the same length with their associated sequences of rules $r(\der{D})=\{\rho_i\}_{i=0}^n$ and $r(\der{D}')=\{\rho'_i\}_{i=0}^n$. A permutation $\sigma:n+1\to n+1$  is called a \emph{consistent permutation} between $\der{D}$ and $\der{D}'$, if for every $i\in n+1$, $\rho_i=\rho'_{\sigma(i)}$ and, moreover, there exists an isomorphism $\xi:\tpro{D} \to \lpro \der{D}' \rpro$ fitting in the following diagrams, where $m_i, m'_i, h_i$ and $h'_i$ are, respectively, the matches and back-matches of $\dder{D}_i$ and $\dder{D}'_i$.
\[\xymatrix{L_i \ar[r]^{m_i} \ar[d]_{m'_{\sigma(i)}}& G_i \ar[r]^{\iota_{G_i}} &\tpro{D} \ar[d]^{\xi} & R_i \ar[r]^{h_i} \ar[d]_{h'_{\sigma(i)}}& G_{i+1} \ar[r]^{\iota_{G_{i+1}}} &\tpro{D} \ar[d]^{\xi} \\G'_{\sigma(i)} \ar[rr]_{\iota_{G'_{\sigma(i)}}}&& \lpro \der{D}' \rpro& G'_{\sigma(i)+1} \ar[rr]_{\iota_{G'_{\sigma(i)+1}}}&& \lpro \der{D}' \rpro}\]
\end{definition}

@@

Take two derivations $\der{D}=\{\dder{D}_{i}\}_{i=0}^n$ and $\der{D}'=\{\dder{D}'_{i}\}_{i=0}^n$ with the same length and between the same $G_0$ and $G_n$. We say that $\der{D}'$ is \emph{obtained by a proper switch from $\der{D}$} if there exists an index $j < n$ such that
	\begin{enumerate}
\item for every $i\notin \{j, j+1\} $, $\dder{D}_i=\dder{D}'_i$;
\item $\dder{D}_j \Updownarrow \dder{D}_{j+1}$;		
\item $\dder{D}'_j\cdot \dder{D}'_{j+1} = S_{u,u'}(\dder{D}, \dder{D}')$.
	\end{enumerate}
	In such a case, we will write $\der{D}\rightsquigarrow_j \der{D}'$ to denote that $\der{D}'$ is obtained by a proper switch between $\dder{D}_j$ and $\dder{D}_{j+1}$.

	We will say that $\der{D}$ is \emph{switch equivalent} to $\der{D}'$, if there exists a sequence, $\{\der{D}_i\}_{i=0}^n$ of derivations such that
	\begin{enumerate}
		\item $\der{D}_0=\der{D}$ and $\der{D}_n=\der{D}'$;
		\item for every $i< n$, $\der{D}_{i+1}$ is obtained by a proper switch from $\der{D}_i$.
	\end{enumerate}

@@

We can now prove some properties of switch equivalence.
\begin{lemma}\todo{Def. 3 della bozza di Andrea }Let $(\X, R)$ be a left-linear DPO-rewriting system. Then the following hold true: 
	\begin{enumerate}
		\item 
		\item 
		\item 
		\item 
	\end{enumerate}
\end{lemma}
\begin{proof}\begin{enumerate}
		\item 
		\item 
		\item 
		\item \qed 
	\end{enumerate}
\end{proof}

@@

\begin{lemma} Let $(\X, R)$ be a left-linear DPO-rewriting system $(\X, R)$. Then the following hold true
	\begin{enumerate}
		\item If $\dder{D}$ and $\dder{D'}$ are two direct derivations such that $\dder{D}\updownarrow \dder{D'}$, then for every filler $(u,u')$ between them, the function
		\[\tau:2\to2 \qquad x \mapsto \begin{cases}
			1 & x=0\\
			0 & x=1
		\end{cases}\]
		defines a consistent permutation between $\dder{D}\cdot \dder{D}'$ and $\sder{D}{D'}$;
		\item if $\der{D}$ and $\der{D}'$ are two switch equivalent derivation, then there exists a consistent permutation between them.
	\end{enumerate}
\end{lemma}
\begin{proof}
	\begin{enumerate}
		\item 
		\item \qed 
	\end{enumerate}
\end{proof}
 This, together with ???
\begin{corollary}
	\todo{unicità}
\end{corollary}
